---
title: #UrbanDataCamp: Using Data to Help Low-Income Youth Find Jobs
author: Keith Hughitt
date: 2015/05/02
output:
    html_document:
        self_contained: false
    md_document:
        variant: markdown_github
---

<div style='font-size: 36px'>#UrbanDataCamp: Using Data to Help Low-Income Youth Find Jobs</div>

```{r knitr_settings, include=FALSE}
library(knitr)
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)
rm(list=ls())    # Clean up any existing variables
```

Overview
========

Questions
---------

1. Can we show that job opportunities for less-privileged youths leads to good
   outcomes later in life? (College enrollment, etc.)

Notes
-----

### General

- "Alumni" 
- "Terminated" - didn't finish

### Factors associated with a positive outcome:

- college enrollment
- college persistence
- alumni

Methods
=======

Load data
---------

```{r load_data, results='asis'}
library(readr)
library(dplyr)

dat = read_csv('input/UA_UI_DataCamp_with_Codebook_4_29.csv')

# preview data
kable(dat[1:5, 1:6])
kable(dat[1:5, 7:12])
kable(dat[1:5, 13:19])
kable(dat[1:5, 20:26])
kable(dat[1:5, 26:33])
```

### Variables

Note: Formatting for metadata is a little bit messed up...

```{r variables, results='asis'}
metadata = read_csv('input/UA_UI_DataCamp_with_Codebook_4_29_metadata.csv')
kable(metadata)

# convert dates to factors so they aren't treated as numeric
dat[['Work Site ZIP Code']]      = factor(dat[['Work Site ZIP Code']])
dat[['Mailing ZIP/Postal Code']] = factor(dat[['Mailing ZIP/Postal Code']])
dat[['HS ZIP Code']]             = factor(dat[['HS ZIP Code']])
```

```{r categorical_variables}
table(dat[['Intern Status']])
```

### Rename variables

For easier typing...

```{r rename_vars}
colnames(dat) = c('datacamp_id', 'academic_year', 'location', 'intern_status',
                  'intern_work_site', 'work_site_zip', 'organization', 'industry',
                  'total_hours_worked', 'dob', 'mailing_zip',
                  'median_household_income', 'labor_force_participation',
                  'unemployment_rate', 'ssi', 'cash_assistance', 'snap',
                  'all_families_below_poverty', 'educational_attainment_ba',
                  'female_headed_households', 'race', 'gender', 'high_school',
                  'title_i_school', 'frl', 'graduation_rate_of_hs',
                  'college_enrollment_of_hs', 'hs_zip_code', 'nsc_data_available',
                  'first_year_college_enrollment', 'first_year_college_type',
                  'first_year_college_state', 'student_quote')

# save with alternative name
write_csv(dat, 'input/UA_UI_DataCamp_with_Codebook_4_29_clean.csv')
```

### Exploratory data analysis

```{r eda}
library(gplots)
```

