---
title: #UrbanDataCamp: Using Data to Help Low-Income Youth Find Jobs
author: Keith Hughitt
date: 2015/05/02
output:
    html_document:
        self_contained: false
    md_document:
        variant: markdown_github
---

<div style='font-size: 36px'>#UrbanDataCamp: Using Data to Help Low-Income Youth Find Jobs</div>

```{r knitr_settings, include=FALSE}
library(knitr)
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)
rm(list=ls())    # Clean up any existing variables
```

Overview
========

Questions
---------

1. Can we show that job opportunities for less-privileged youths leads to good
   outcomes later in life? (College enrollment, etc.)

Notes
-----

### General

- "Alumni" 
- "Terminated" - didn't finish

### Factors associated with a positive outcome:

- college enrollment
- college persistence
- alumni

Methods
=======

Load data
---------

```{r load_data, results='asis'}
library(readr)
library(dplyr)

dat = read_csv('input/UA_UI_DataCamp_with_Codebook_4_29.csv')

# preview data
kable(dat[1:5, 1:6])
kable(dat[1:5, 7:12])
kable(dat[1:5, 13:19])
kable(dat[1:5, 20:26])
kable(dat[1:5, 26:33])
```

### Variables

Note: Formatting for metadata is a little bit messed up...

```{r variables, results='asis'}
metadata = read_csv('input/UA_UI_DataCamp_with_Codebook_4_29_metadata.csv')
kable(metadata)

# convert dates to factors so they aren't treated as numeric
dat[['Work Site ZIP Code']]      = factor(dat[['Work Site ZIP Code']])
dat[['Mailing ZIP/Postal Code']] = factor(dat[['Mailing ZIP/Postal Code']])
dat[['HS ZIP Code']]             = factor(dat[['HS ZIP Code']])
```

```{r categorical_variables}
table(dat[['Intern Status']])
```

### Rename variables

For easier typing...

```{r rename_vars}
colnames(dat) = c('datacamp_id', 'academic_year', 'location', 'intern_status',
                  'intern_work_site', 'work_site_zip', 'organization', 'industry',
                  'total_hours_worked', 'dob', 'mailing_zip',
                  'median_household_income', 'labor_force_participation',
                  'unemployment_rate', 'ssi', 'cash_assistance', 'snap',
                  'all_families_below_poverty', 'educational_attainment_ba',
                  'female_headed_households', 'race', 'gender', 'high_school',
                  'title_i_school', 'frl', 'graduation_rate_of_hs',
                  'college_enrollment_of_hs', 'hs_zip_code', 'nsc_data_available',
                  'first_year_college_enrollment', 'first_year_college_type',
                  'first_year_college_state', 'student_quote')

# save with alternative name
write_csv(dat, 'input/UA_UI_DataCamp_with_Codebook_4_29_clean.csv')
```

### Exploratory data analysis

#### Variable relationships

```{r eda}
library(gplots)
library(bpca)
library(RColorBrewer)

# create short version of data with a subset of the fields and not including
# the current (2014-2015) interns who haven't yet finished; remove rows with
# duplicated ids
duplicate_ids = names(table(dat$datacamp_id)[table(dat$datacamp_id) > 1])
dat_short = dat %>%
    filter(academic_year != '2014 - 2015' & 
           industry != "" & 
           !datacamp_id %in% duplicate_ids) %>%
    select(datacamp_id, intern_status, industry, total_hours_worked,
           median_household_income, labor_force_participation,
           unemployment_rate, ssi, cash_assistance, snap,
           all_families_below_poverty, educational_attainment_ba,
           female_headed_households, graduation_rate_of_hs) %>%
    mutate(intern_status=ifelse(intern_status %in% 
                                c("Alumni", "Summer Program Alumni"),
                                "Alumni", "Terminated"))
    
dat_short_complete = dat_short[complete.cases(dat_short),]

ids = dat_short_complete$datacamp_id

dat_short_complete = dat_short_complete %>% select(-datacamp_id)

# Heatmap
heatmap.2(cor(as.matrix(dat_short_complete %>% select(-intern_status, -industry))),
          margins=c(26, 26), trace='none')

# group by location
dat$location_colors = brewer.pal(length(unique(dat$location)), "Set1")[
    as.numeric(as.factor((as.character(dat$location))))]
location_colors_complete = dat$location_colors[dat$datacamp_id %in% ids]

# biplot
plot(bpca(dat_short_complete %>% select(-intern_status, -industry),
          var.color=location_colors_complete, scale=TRUE))
```

#### Students

```{r student_vis}
# rescale data
rescale = function(x) {
    (x - min(x)) / max(x)
}

dat_scaled = dat_short_complete

dat_scaled$median_household_income  = rescale(dat_scaled$median_household_income)
dat_scaled$total_hours_worked       = rescale(dat_scaled$total_hours_worked)
dat_scaled$ssi                      = rescale(dat_scaled$ssi)
dat_scaled$unemployment_rate        = rescale(dat_scaled$unemployment_rate)
dat_scaled$female_headed_households = rescale(dat_scaled$female_headed_households)
dat_scaled$cash_assistance          = rescale(dat_scaled$cash_assistance)
dat_scaled$snap                     = rescale(dat_scaled$snap)
dat_scaled$all_families_below_poverty = rescale(dat_scaled$all_families_below_poverty)

# numeric fields only
#mat = dat_short_complete %>% select(-intern_status)
#mat_scaled = scale(dat_short_complete)
heatmap.2(as.matrix(dat_scaled %>% select(-intern_status, -industry)),
          margins=c(26, 26), trace='none',
          RowSideColors=location_colors_complete)
```

#### Median household income vs. Hours worked 

```{r household_income_vs_intern_status}
library(ggplot2)
ggplot(dat_short_complete, aes(total_hours_worked, median_household_income,
                               group=intern_status), color=intern_status) +
    geom_point(aes(color=intern_status))
```

#### Intern status vs. Industry

```{r intern_status_vs_industry}
dat_short_complete %>% group_by(industry) %>% ggplot(aes(intern_status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    geom_bar() +
    facet_wrap(~industry)

#ggplot(dat_short_complete, aes(intern_status, total_hours_worked,
#                               group=industry), color=industry) +
#    geom_point(aes(color=industry))
```

Session info
============

```{r system_info}
date()
sessionInfo()
```

